Locally-Oriented Programming:\\ A Simple Programming Model for Stencil-Based Computations on Multi-Level Distributed Memory Architectures

### 2015-02-07 papers/LOPe/local.tex
Added sections on transformations and conclusions.
Paper lenght shortening begun.

\usepackage{amssymb}
\setcounter{tocdepth}{3}
\usepackage{graphicx}
\usepackage{todonotes}
\usepackage{url}
\urldef{\mailsa}\path|rasmus@uoregon.edu, mjsottile@gmail.com,|
\urldef{\mailsb}\path|dangle@ucar.edu, soren.rasmussen@aggiemail.usu.edu|
\newcommand{\keywords}[1]{\par\addvspace\baselineskip
\noindent\keywordname\enspace\ignorespaces#1}

@@

% the affiliations are given next; don't give your e-mail address
% unless you accept that it will be published
\institute{441 McKenzie Hall,\\
5246 University of Oregon, Eugene, Oregon, USA\\
\mailsa\\
\mailsb\\
\url{https://openfortran.uoregon.edu}}

@@

Emerging hybrid accelerator architectures for high performance computing are often suited for the
use of a data-parallel programming model.  Unfortunately, programmers for these emerging
architectures face a steep learning curve that frequently requires learning a new language, e.g.,
OpenCL or CUDA, or employing OpenMP compiler directives.  Furthermore, the distributed (and
frequently multi-level) nature of the memory organization of clusters of these architectures
provides an additional level of complexity for the programmer.  This paper presents preliminary work
examining how programming with a local orientation can be employed to provide simple access to
accelerator architectures.  A locally-oriented programming model is especially useful for the
solution of algorithms requiring the application of a convolution kernel.  In this programming
model, a programmer codes the algorithm by modifying \emph{only a single array element} (called the
local element), but has read-only access to a small sub-array surrounding the local element so that
a stencil can be applied.  We demonstrate how a locally-oriented programming model can be adopted as
an embedded, domain-specific language using source-to-source program transformations.

@@

%\input{motivation}
\input{intro}
\input{programming-model}
\input{coarray-extensions}

>>>>

\usepackage{amssymb}
\setcounter{tocdepth}{3}
\usepackage{graphicx}
\usepackage{todonotes}
\usepackage{url}
%%\urldef{\mailsa}\path|rasmus@uoregon.edu, mjsottile@gmail.com,|
\urldef{\mailsa}\path|rasmus@uoregon.edu|
\urldef{\mailsb}\path|dangle@ucar.edu, soren.rasmussen@aggiemail.usu.edu|
\newcommand{\keywords}[1]{\par\addvspace\baselineskip
\noindent\keywordname\enspace\ignorespaces#1}

@@

% the affiliations are given next; don't give your e-mail address
% unless you accept that it will be published
%%\institute{441 McKenzie Hall,\\
%%5246 University of Oregon, Eugene, Oregon, USA\\
%%\mailsa\\
%%\mailsb\\
%%\url{https://openfortran.uoregon.edu}}

\institute{University of Oregon, Eugene, Oregon, USA\\
\mailsa
}

@@

Emerging hybrid accelerator architectures for high performance computing are often suited for the
use of a data-parallel programming model.  Unfortunately, programmers of these
architectures face a steep learning curve that frequently requires learning a new language (e.g.,
OpenCL). Furthermore, the distributed (and
frequently multi-level) nature of the memory organization of clusters of these machines
provides an additional level of complexity.  This paper presents preliminary work
examining how programming with a local orientation can be employed to provide simpler access to
accelerator architectures.  A locally-oriented programming model is especially useful for the
solution of algorithms requiring the application of a stencil or convolution kernel.  In this programming
model, a programmer codes the algorithm by modifying \emph{only a single array element} (called the
local element), but has read-only access to a small sub-array surrounding the local element.
We demonstrate how a locally-oriented programming model can be adopted as
a language extension using source-to-source program transformations.

@@

%\input{motivation}
\input{intro}
\input{programming-model}
\input{coarray-extensions}
\input{transformations}
\input{conclusions}
